---
title: "Demographic Breakdowns by Experimental Phase: Job Applicants Experiment"
output: rmarkdown::github_document
date: "2024-01-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

# Loading survey data

* Includes warmup, iterative batch phases, and validation

```{r}
df <- readRDS("../../02_output/job_applicants_data_clean.RDS")

df <- df %>%
  mutate(batch_type = factor(batch_type,
    levels = c("Warmup", "Iterative Batch Phase: Max", "Iterative Batch Phase: Min", "Validation"),
    ordered = TRUE
  ))

# look at sample counts by batch id and batch type
df %>%
  group_by(batch_type, batch_id) %>%
  summarize(n = n())

df %>%
  group_by(batch_type) %>%
  summarize(n = n())
```

## Examining female and hispanicity

* There are some differences in the percent female between the different phases (around 10 percentage points)

```{r}
## look at female and hispanicity
df %>%
  group_by(batch_type) %>%
  summarize(across(c("female", "hispanic"), .fns = lst(
    n = ~ sum(.),
    per = ~ mean(.)
  )))
```

## Examining differences in ages

* Distribution seems similar across phases

```{r}
# age
df %>%
  group_by(batch_type, age) %>%
  summarize(n = n()) %>%
  group_by(batch_type) %>%
  mutate(per = n / sum(n)) %>%
  arrange(age)

df %>%
  # remove outlier
  filter(age < 100) %>%
  ggplot(aes(age, after_stat(density), fill = batch_type)) +
  geom_histogram() +
  facet_wrap(~batch_type)
```

## Examining differences in race

* Distribution also is similar across batches

```{r}
# race
df %>%
  group_by(batch_type) %>%
  group_by(race) %>%
  summarize(n = n()) %>%
  ungroup() %>%
  mutate(per = n / sum(n)) %>%
  arrange(desc(per))
```

```{r, fig.height=7, fig.width=8}
df %>%
  group_by(batch_type, race) %>%
  summarize(n = n()) %>%
  group_by(batch_type) %>%
  mutate(per = n / sum(n)) %>%
  arrange(desc(per)) %>%
  ggplot(aes(reorder(race, desc(per)), per, fill = batch_type)) +
  geom_col() +
  facet_wrap(~batch_type) +
  labs(
    x = "Race",
    y = "Percent of total"
  ) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```



